---
title: "p8105_hw5_xz3173"
author: "Xue Zhang"
date: "2023-11-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(broom)
```

# Problem 1

```{r}
# Load homicide dataset
homicide_df = 
  read_csv("data/homicide-data.csv") |>
  janitor::clean_names()

homicide_df
```

Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
# create a city_state variable
city_summary = homicide_df |>
  mutate(city_state = str_c(city, state, sep = ", ")) |>

# summarize within cities to obtain the ttotal number of homicides and the number of unsolved homicides
  group_by(city_state) |>
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  )
```

For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
# filter data for Baltimore, MD
baltimore_df = city_summary |>
  filter(city_state == "Baltimore, MD") |>

# using prop.test to estimate the proportion of unsolved homicides
  summarize(
    test_result = list(
      prop.test(
        x = unsolved_homicides,
        n = total_homicides) |>
        tidy()
    )
  ) |>
  pull(test_result)

# check the results
baltimore_df
```